Evolution of pre-copulatory and post-copulatory inbreeding behaviour: a computational approach

### 2016-09-15 ms_pre_post.Rmd
Add abstract, a bit more Discussion

@@

Figure 2, which we previously contrasted with Figure 1 to test whether or not evolution of inbreeding fertilisation strategy facilitates selection for polyandry, shows evolution of tendency for polyandry and inbreeding mating strategy when inbreeding fertilisation strategy is fixed such that the sire of a focal female's offspring is randomly assigned from her mates according to a fair raffle. In all cost combinations of Figure 2 (panels A-D), tendency for polyandry allele values decreased over generations, reflecting evolution of monandry. However, values of inbreeding mating strategy alleles did not greatly differ over generations when inbreeding fertilisation strategy was fixed, meaning that the effect of fixed inbreeding fertilisation strategy on increasing selection for inbreeding avoidance through mate choice was weak or nelgigible. A weak effect is discernable in the contrast between Figure 2B and Figure 1B, in which the grand mean inbreeding mating strategy allele value is $-0.627$ when inbreeding fertilisation strategy is not fixed (Figure 1B) versus $-1.036$ when it is fixed (Figure 2B) in generation 10000. Because females have 20 alleles with additive effects on inbreeding mating strategy, an average female is therefore expected to have mean inbreeding mating strategy phenotype values of $M_{p} = -0.627 \times 20 = -12.54$ and $M_{p} = -1.036 \times 20 = -20.72$ for not fixed versus fixed inbreeding fertilisation strategy, respectively. These values would cause an outbred focal female to be $2.135$ and $4.180$ times more likely to choose a non-relative than a full sibling as a mate given not fixed fersus fixed inbreeding fertilisation strategy, respectively. Therefore, when females were unable to bias fertilisation to avoid inbreeding given severe inbreeding depression ($\beta=3$), selection for evolution of pre-copulatory inbreeding avoidance was strengthened.

@@

>>>>

Inbreeding avoidance has been observed through pre-copulatory and post-copulatory mechanisms in empirical systems as a consequence of inbreeding depression. However, no models compare evolution of pre-copulatory versus post-copulatory inbreeding avoidance jointly, and given biologically relevant costs of, and constraints on, inbreeding strategies. Critically, evolution of post-copulatory inbreeding avoidance is restricted to females that mate multiply, potentially leading to complex evolutionary dynamics among inbreeding strategies and polyandry. Accordingly, use individual-based modelling to track evolution of polyandry and inbreeding strategies under multiple costs and constraints to address four questions: (1) Will evolution of post-copulatory inbreeding avoidance strengthen selection for polyandry? (2) Will selection for one costly inbreeding strategy be negligible given a non-costly complementary strategy? (3) Will constraints on evolution of one inbreeding strategy increase selection on an alternative strategy? And (4) will selection for one inbreeding strategy be negligible given existence of an already adaptive alternative strategy? We show that evolution of polyandry is greatly facilitated by evolution of post-copulatory inbreeding avoidance. Costs and constraints on pre-copulatory inbreeding avoidance generally affected evolution of post-copulatory inbreeding avoidance, but not vice versa. Our model thereby introduces a novel framework for predicting evolution of both pre-copulatory and post-copulatory inbreeding strategies in empirical systems.

@@

Figure 2, which we previously contrasted with Figure 1 to test whether or not evolution of inbreeding fertilisation strategy facilitates selection for polyandry, shows evolution of tendency for polyandry and inbreeding mating strategy when inbreeding fertilisation strategy is fixed such that the sire of a focal female's offspring is randomly assigned from her mates according to a fair raffle. In all cost combinations of Figure 2 (panels A-D), tendency for polyandry allele values decreased over generations, reflecting evolution of monandry. However, values of inbreeding mating strategy alleles did not greatly differ over generations when inbreeding fertilisation strategy was fixed, meaning that the effect of fixed inbreeding fertilisation strategy on increasing selection for inbreeding avoidance through mate choice was weak or nelgigible. A weak effect is discernable in the contrast between Figure 2B and Figure 1B, in which the grand mean inbreeding mating strategy allele value is $-0.627$ when inbreeding fertilisation strategy is not fixed (Figure 1B) versus $-1.036$ when it is fixed (Figure 2B) in generation 10000. Because females have 20 alleles with additive effects on inbreeding mating strategy, an average female is therefore expected to have mean inbreeding mating strategy phenotype values of $M_{p} = -0.627 \times 20 = -12.54$ and $M_{p} = -1.036 \times 20 = -20.72$ for not fixed versus fixed inbreeding fertilisation strategy, respectively. These values would cause an outbred focal female to be $2.135$ and $4.180$ times more likely to choose a non-relative than a full sibling as a mate given not fixed versus fixed inbreeding fertilisation strategy, respectively. Therefore, when females were unable to bias fertilisation to avoid inbreeding given severe inbreeding depression ($\beta=3$), selection for evolution of pre-copulatory inbreeding avoidance was strengthened.

@@

<!--- Also: note, pre-copulatory but not post-copulatory inbreeding avoidance can occur for males --->
